---
title: "freeze_vs_cold"
author: "Lauren Gill"
date: "31/07/2023"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
 This is an analysis of the data for freeze vs cold
 
 Load libraries
```{r}
library(tidyverse)
library(officer)
library(rvg)
```

Read in dataset
```{r}
cold <- read_csv("coldvsfreeze.csv")
view(cold)
```

How to display survival and freezing ....
Trying a stacked bar chart showing time spent frozen and time spent at cold exposure
First create data that will suit this type of plot
```{r}
cold$slug_number <- as.factor(cold$slug_number)
cold$survival <- as.factor(cold$survival)
cold$frozen <- as.factor(cold$frozen)

stacked_bar <- cold %>%
  dplyr::select(survival, cold_exposure_length, time_spent_frozen, slug_number) %>%
  mutate(only_cold = cold_exposure_length - time_spent_frozen) %>%
  pivot_longer(cols = c(time_spent_frozen, only_cold), 
               names_to = "state", 
               values_to = "time")

stacked_bar
```

Testing this with logistic regression
```{r}
cold.glm <- glm(survival ~ weight + cold_exposure_length + frozen, data = cold, family="binomial")
summary(cold.glm)
anova(cold.glm,test="Chisq")
```

Now creating a stacked bar plot
```{r}
level_order <- c("time_spent_frozen", "only_cold")

cold_bar <- ggplot(data = stacked_bar, aes(x = time, y = slug_number, fill = factor(state, levels = level_order))) + 
    geom_bar(stat='identity') +
  facet_grid(survival ~ .) +
  theme_bw()

cold_bar
```

Exporting it to power point to make it look pretty
https://rpubs.com/techanswers88/VectorGraphicsChartsUsingGGPLOT
```{r}
#first export to vector object
myplot <-  dml(ggobj = cold_bar
              ,bg = "white"
              ,pointsize = 12
              ,editable = TRUE)
doc <- read_pptx()


doc <- add_slide(doc, "Title and Content", "Office Theme")
doc <- ph_with(doc, myplot, location = ph_location_fullsize())

# If you want to create multiple charts then create your second chart and call it myplot2 

#doc <- add_slide(doc, "Title and Content", "Office Theme")
#doc <- ph_with(doc, myplot, location = ph_location_fullsize())

fileout <- "mygraph.pptx"
print(doc, target = fileout)

```


Trying to make a mosaic plot
```{r}
cold %>%
  group_by(frozen, survival) %>%
  summarise(number = length(weight))

%>%
  ggplot(aes(x = frozen, y = number, fill = survival )) +
  geom_col()
```

