---
title: "NMR analysis"
author: "Lauren Gill"
date: "10/01/2024"
output: pdf_document
---

Load in libraries

```{r}
library(tidyverse)
library(mdatools)
library(ggpubr)
library(car)
library(emmeans)
library(cowplot)
library(multcompView)
library(agricolae)
library(openxlsx)
```

Load in data The data is the metabolites already normalized by wetweight

```{r}
concentrations <- read_csv("../data/concentrations_names.csv")
slug_weights <- read_csv("../data/slug_weights.csv")
```

## Transform data

First making the data into um intstead of uM
The slugs were placed in 500uL (500/10^6 L) of acetonitrile so concentrations are relative to this

Concentrations in the original table are in uM
uM = umol/L 
umol = uM * L
```{r}
500/(10^6)
## renaming the sample column in slug_weights
slug_weights <- slug_weights %>% 
  rename(sample = slug_number)

nmr <- right_join(slug_weights, concentrations, by = "sample")%>%
  #converting to umoles (see notes above)
  mutate(across(`1-Methylhistidine`:`Dimethyl sulfone`,
                ~ .*(500/(10^6)))) %>%
  #dividing each conc in each row by the dry weight (in g)
  mutate(across(`1-Methylhistidine`:`Dimethyl sulfone`,
                ~ ./wet_weight_g)) %>%
  #removing extrenuous colomns (for metaboanalyst) 
  dplyr::select(-c(wet_weight_g, dry_weight_g))

#write_csv(nmr_wet, "new_wet_conc_moles.csv")
```
Final units: umol metabolite/g ww

```{r}
nmr_transformed <- nmr %>%
  #sqrt tranforming the data
  mutate(across(!c(sample, group), sqrt)) %>%
  #deleting the first two columns because "prep.autoscale" doesnt use them
  dplyr::select(!c(sample, group))

#Now autoscaling the data
nmr_scaled <- prep.autoscale(nmr_transformed, center = TRUE, scale = TRUE)

#Adding back in the first two columns
first_columns <- nmr %>%
  dplyr::select(c(sample, group))
nmr_normal <- cbind(first_columns, nmr_scaled)
```

Check out this transformed data

```{r}

#Original NMR data (big variances)
nmr_original <- nmr %>%
  dplyr::select(!c(sample, group))
boxplot(nmr_original)

#the transformed data, looks a lot better!
boxplot(nmr_scaled)
```

## ANOVA

Trying a two way anova (<https://haowang47.github.io/files/2023-01-31-2wayANOVA.html>) 

Lets make a function
Input = dataname, name of metabolite
```{r}
metaboanova <- function(data, metabolite){
#first making a data table with just metabo concentrations

glut <- data %>%
  dplyr::select(sample, group, {{metabolite}}) %>%
  rename(conc = {{metabolite}}) %>%
  #seperating SD20 into SD and 20 
  tidyr::separate(group,
                  into = c("day_length", "temp"),
                  sep = "(?<=[A-Za-z])(?=[0-9])")

glut_anova <- lm(conc ~ day_length*temp, data = glut, 
                   contrasts = list(temp="contr.sum", day_length="contr.sum"))


aov(glut_anova)
}
```


```{r}
metaboletters <- function(data, metabolite){
#first making a data table with just metabo concentrations

glut <- data %>%
  dplyr::select(sample, group, {{metabolite}}) %>%
  rename(conc = {{metabolite}}) %>%
  #seperating SD20 into SD and 20 
  tidyr::separate(group,
           into = c("day_length", "temp"), 
           sep = "(?<=[A-Za-z])(?=[0-9])")

glut_anova <- lm(conc ~ day_length*temp, data = glut, 
                   contrasts = list(temp="contr.sum", day_length="contr.sum"))


multcompLetters4(glut_anova, TukeyHSD(aov(glut_anova)), reversed = TRUE)
}

```

Getting letters to put on the plot
```{r}
th_letter <- metaboletters(nmr_normal, `L-Threonine`)
th_letter_1 <- th_letter$`day_length:temp`
th_letter_1

metaboletters(nmr_normal, `L-Glutamine`)
metaboletters(nmr_normal, `Formate`)
```



```{r}
th <- metaboanova(nmr_normal, `L-Threonine`)
summary(th)
th_tukey <- TukeyHSD(th)
letters <- multcompLetters4(th, TukeyHSD(aov(th)), reversed = TRUE)
th_tukey

glut <- metaboanova(nmr_normal, `L-Glutamine`)
summary(glut)
form <- metaboanova(nmr_normal, `Formate`)
summary(form)
```

Trying to add p values
```{r}

nmr_summarized <-  nmr %>%
  dplyr::select(sample, group, Formate, `L-Threonine`, `L-Glutamine`)%>%
  pivot_longer(!c(sample, group), names_to = "metabolite", values_to = "conc") %>%
  group_by(group) %>% 
  summarize(max_conc=max(conc))

hsd <- HSD.test(th, "day_length", group=T)

```




### Mean Comparisons

Checking simple effects
#### Glutamine
```{r}
#first making a data table with just metabo concentrations
glut <- nmr_normal %>%
  dplyr::select(sample, group, `L-Glutamine`) %>%
  rename(conc = `L-Glutamine`) %>%
  #seperating SD20 into SD and 20 
  tidyr::separate(group,
           into = c("day_length", "temp"), 
           sep = "(?<=[A-Za-z])(?=[0-9])")

glut_anova <- lm(conc ~ day_length*temp, data = glut, 
                   contrasts = list(temp="contr.sum",
                                    day_length="contr.sum"))

#by day length
tempbydaylength <- emmeans(glut_anova, ~temp|day_length)
tempbydaylength

pairs(tempbydaylength) #signficant!

#by temp
daylengthbytemp <- emmeans(glut_anova, ~day_length|temp)
daylengthbytemp

pairs(daylengthbytemp) #nothing significant!
```


### Model fit

Does the Model fit the data?? QQplot - looks good

```{r}
ggplot(glut_anova, aes(sample = rstandard(glut_anova))) + geom_qq() + stat_qq_line()+ylab("Studentized Residuals Qunatiles") + xlab("Normal Qunatiles")
```

### Plotting

```{r}
glut_graph <- ggplot(glut, aes(x = temp, y = conc)) +
  geom_boxplot() +
  geom_jitter() +
  facet_wrap(~day_length) +
  theme_cowplot() +
  stat_compare_means(label =  "p.signif", label.x = 1.5)

```

## Making a Metabo Graphing Function

now make a function for graphing by metabolite

```{r}

metabograph <- function(data, metabolite, x, wrap){
  
  data %>%
  dplyr::select(sample, group, {{metabolite}}) %>%
  rename(conc = {{metabolite}}) %>%
  #seperating SD20 into SD and 20 
  separate(group,
           into = c("day_length", "temp"), 
           sep = "(?<=[A-Za-z])(?=[0-9])") %>%
    ggplot(aes(x = {{x}}, y =conc)) +
  geom_boxplot() +
  geom_jitter(width = 0.28) +
  facet_wrap(enquo(wrap)) +
    theme_cowplot() +
    labs(y = expression(paste("Normalized Metabolite Concentration (µM m",g^-1,"dry mass)")))+
    stat_compare_means(label =  "p.signif", label.x = 1.5)

}

metabograph(nmr_normal, `L-Glutamine`, temp, day_length)
#ggsave("glutamine.pdf")
```

Lets try this function on threonine

```{r}
p1 <- metabograph(nmr_normal, `L-Threonine`, day_length, temp)
#ggsave("threonine.pdf")

```

Lets try this function on formate

```{r}
metabograph(nmr_normal, Formate, temp, day_length)
#ggsave("fomate.pdf")
```

Now making a new graph that has a multiplot set up
This time I am using the un-transformed data

```{r fig.height=5.5, fig.width=5}

 nmr %>%
  dplyr::select(sample, group, Formate, `L-Threonine`, `L-Glutamine`)%>%
  pivot_longer(!c(sample, group), names_to = "metabolite", values_to = "conc") %>%
    ggplot(aes(x = group, y =conc, fill = group)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.3) +
  geom_jitter(width = 0.25, alpha = 0.3) +
  scale_x_discrete(limits = c("LD20", "LD15", "SD20", "SD15"))+
  
  facet_wrap(vars(metabolite), nrow = 3, scales = "free", strip.position = "top") +
    theme_cowplot() +
  #get rid of legend
theme(legend.position="none")+
    labs(y = expression(paste("Metabolite Concentration (µmol ",g^-1,"wet mass)")),
         x = "Group")

#getting n values
nmr %>%
  group_by(group) %>%
  summarize(count = n())

#ggsave("all_metabolites_plot_updatedoct2025.pdf")
```


Table for results
```{r}

#read in file with metabolite groups
mgroup <- read_csv("../data/metabolite_groups.csv")
mgroup_long <- mgroup %>%
  pivot_longer(cols = everything(),
               names_to = "category", 
               values_to = "metabolite",
               values_drop_na = TRUE)

table_for_results <- nmr %>%
  dplyr::select(!c(group)) %>%
  pivot_longer(!sample, names_to = "metabolite", values_to = "concentrations") %>%
  group_by(metabolite) %>%
  summarise(N    = length(concentrations),
                     mean = mean(concentrations, na.rm = TRUE),
                     sd   = sd(concentrations, na.rm = TRUE),
                     std   = sd / sqrt(N)) %>%
  #rounding
  mutate_at(vars(-metabolite), funs(round(., 2))) %>%
  inner_join(mgroup_long,
             by = "metabolite") %>%
#paste together mean and std and metabolite
  mutate(all = paste("[",mean,"±",std, "]"),
         #get rid of white spaces
         all = gsub(" ", "", all),
         #paste the metabolite name again
         all2 = paste(metabolite, all))%>%
   #get rid of N and sd
  dplyr::select(all2, category) %>%
  ungroup() %>%
mutate(row_id = row_number()) %>%
  # Pivot wider while using 'row_id' to avoid aggregation issues
  pivot_wider(
    names_from = category,
    values_from = all2,
    id_cols = row_id) %>%
  dplyr::select(-c(row_id))

#write.xlsx(table_for_results, 'table_for_results_updated.xlsx')
```

