---
title: "slug_survival"
author: "Lauren Gill"
date: "24/07/2023"
output: html_document
---

Checking out slug survival

# Load libraries

```{r}
library(tidyverse)
library(cowplot)
library(multcomp)
library(MASS)
library(scales)
library(multcompView)
library(emmeans)
library(sciplot)
```

# Read in dataset

```{r}
survival_original <- read_csv("../data/survival_original.csv")
slug_weights <- read.csv("../data/slug_weights.csv")

```

# Body weight

```{r}
body_weight <- survival_original %>%
  filter(minutes_frozen == 0) %>%
  summarize(mean_weight = mean(weight),
            sd_weight = sd(weight),
            count = n(),
            se=sd_weight/sqrt(count),
            max = max(weight),
            min = min(weight)) %>%
  drop_na()
```

Compare bodyweight by survival in each category

```{r}
survival_body_weight <- survival_original %>%
  filter(minutes_frozen == 0) %>%
  group_by(treatment, survival) %>%
  summarize(mean_weight = mean(weight),
            sd_weight = sd(weight),
            count = n(),
            se=sd_weight/sqrt(count), 
            upper_limit=mean_weight+se, 
            lower_limit=mean_weight-se 
  ) %>%
  drop_na()

survival_body_weight$survival <- as.factor(survival_body_weight$survival)

survival_body_weight %>%
  ggplot(aes(x = survival, y = mean_weight)) +
  geom_point(size = 3)+
  geom_errorbar(aes(ymin=lower_limit, ymax=upper_limit))+
  facet_wrap(vars(treatment)) +
  theme_bw()

str(survival_body_weight$survival)

p2 <- survival_original %>%
  ggplot(aes(x = weight, y = averageSCP)) +
  geom_point() +
  geom_smooth(method = 'lm') +
  theme_classic() +
  labs(x = "Body Mass (g)",
       y = "Supercooling Point (°C)") +
    scale_x_continuous(limits = c(0.1, 1.1), breaks = pretty_breaks())

p2


```

# Survival

Look at survival proportions for all slugs Standard Error of the proportion (where p is the sample proportion) = √p̂(1-p̂) / n

```{r}

survival_by_treatment <- survival_original %>%
  filter(minutes_frozen == 0) %>%
  group_by(treatment) %>%
  drop_na(survival) %>%
  summarize(prop_surv = mean(survival, na.rm = TRUE),
            count = n(),
            se=sqrt((prop_surv*(1-prop_surv))/count), 
            upper_limit=prop_surv+se, 
            lower_limit=prop_surv-se) %>%
  arrange((prop_surv))


```

Now running a chisq test on the data

```{r}

survival_stats <- survival_original %>%
  filter(minutes_frozen == 0) %>%
  mutate(acclimation_day_length = as_factor(acclimation_day_length),
         acclimation_temp = as_factor(acclimation_temp))

slug.survival.glm <- glm(survival~ acclimation_day_length*acclimation_temp + weight,family="binomial",data=survival_stats)
anova(slug.survival.glm,test="Chisq")


```

Running a Graph

```{r}

level_order <- c("LD20", "LD15", "SD20", "SD15")
p1 <- survival_by_treatment %>%
  ggplot(aes(x = factor(treatment, level = level_order), y = prop_surv)) +
  geom_point(size = 3)+
  geom_errorbar(aes(ymin=lower_limit, ymax=upper_limit), width = 0.5)+
  xlab("Treatment")+
  ylab("Survival Proportion")+
  scale_y_continuous(limits = c(0, 0.6), breaks = pretty_breaks())+
  theme_classic(base_size = 18) 

p1

```

# Super cooling Point

Now looking at the average SCP for the slugs

```{r}
scp_by_treatment <- survival_original %>%
  group_by(treatment) %>%
  summarize(mean_scp = mean(averageSCP, na.rm = TRUE),
            count = n(),
            sd_scp=sd(averageSCP, na.rm = TRUE), 
            se=sd_scp/sqrt(count), 
            upper_limit=mean_scp+se, 
            lower_limit=mean_scp-se)

scp.glm <- glm(averageSCP~weight+acclimation_day_length*acclimation_temp,data=survival_stats)
summary(aov(scp.glm))
scp.anova <- aov(averageSCP~weight+acclimation_day_length*acclimation_temp,data=survival_stats)

summary(scp.anova)
```

Making a plot of scp by treatment

```{r}
p2 <- scp_by_treatment %>%
  ggplot(aes(x = factor(treatment, level = level_order), y = mean_scp)) +
  geom_point(size = 3)+
  geom_errorbar(aes(ymin=lower_limit, ymax=upper_limit), width = 0.5)+
  xlab("Treatment")+
  theme_bw()

p2

```

Now making a line plot

```{r}
#relationship between weight and average scp?
plot(survival_original$weight, survival_original$averageSCP)

#lineplot


lineplot.CI(acclimation_temp,averageSCP,acclimation_day_length,
            data=survival_original,
            lwd = 2,
            cex = 1,
            col = c("light blue", "dark blue"),
            leg.lab = c("Short Day", "Long Day"),
            xlab = "Acclimation Temperature (°C)",
            ylab = "Supercooling Point (°C)",
            las = 1,
            ylim = c(-8, -2),
            cex.leg = 1.1,
            x.leg = 1.8,
            cex.axis = 1.2) 
```

Now stacking these graphs upon eachother

```{r}
plot_grid(
  p2, p1,
  labels = "AUTO", ncol = 1,
  rel_heights = c(0.4, 0.7)
)
```

# Body Water Content

Looking at wet weight vs dry weight

```{r}
slug_weights_deets <- slug_weights %>%
  #photoperiod
   mutate(
    photoperiod = case_when(
      str_sub(slug_number, 1, 1) == 'A' ~ 'LD',
      str_sub(slug_number, 1, 1) == 'B' ~ 'SD',
      str_sub(slug_number, 1, 1) == 'C' ~ 'LD',
      str_sub(slug_number, 1, 1) == 'D' ~ 'SD'
    )
  ) %>%
  #temperature
     mutate(
    temperature = case_when(
      str_sub(slug_number, 1, 1) == 'A' ~ '20',
      str_sub(slug_number, 1, 1) == 'B' ~ '20',
      str_sub(slug_number, 1, 1) == 'C' ~ '15',
      str_sub(slug_number, 1, 1) == 'D' ~ '15'
    )
  ) %>%
  #difference between wet and dry
  mutate(dryoverwet = dry_weight_g/wet_weight_g) %>%
  #wet percentage
  mutate(wetpercent = (wet_weight_g -dry_weight_g)/wet_weight_g)

slug_weights_grouped <- slug_weights_deets %>%
  group_by(photoperiod, temperature) %>%
  summarise(N    = length(wetpercent),
                     mean = mean(wetpercent, na.rm = TRUE),
                     sd   = sd(wetpercent, na.rm = TRUE),
                     std   = sd / sqrt(N))

slug_weights_grouped_temp <- slug_weights_deets %>%
  group_by(temperature) %>%
  summarise(N    = length(wetpercent),
                     mean = mean(wetpercent, na.rm = TRUE),
                     sd   = sd(wetpercent, na.rm = TRUE),
                     std   = sd / sqrt(N))

#graph
weight_graph <- slug_weights_grouped%>%
  ggplot(aes(x = temperature, y = mean, fill = photoperiod)) +
  geom_point(size=3, 
             position = position_dodge(width = 1),
             aes(colour = photoperiod)) +
  geom_errorbar(position = position_dodge(width = 1),
                linewidth=0.4,
                width = 0.5,
                aes(ymin=mean-std, ymax=mean+std))+
  #adding in jitter points
    geom_point(data = slug_weights_deets, 
               aes(colour = photoperiod, x= temperature, y = wetpercent),
               position = position_dodge(width = 1), 
               size = 1.3, alpha = .3,
               show.legend = FALSE)+
  labs(y = "Dry weight/Wet Weight") +
  theme_classic()

weight_graph

```

Stats

```{r}
lm_weight <- lm(wetpercent ~ photoperiod * temperature, data = slug_weights_deets)
anova(lm_weight)
```
